\documentclass[a4paper]{exam}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{array}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{titling}

\newcolumntype{C}{>{$}c<{$}} % math-mode version of "c" column type

\runningheader{CS/MATH 113}{WC04: Logical Inference}{\theauthor}
\runningheadrule
\runningfootrule
\runningfooter{}{Page \thepage\ of \numpages}{}

\printanswers

\title{Weekly Challenge 04: Logical Inference\\CS/MATH 113 Discrete Mathematics}
\author{team-name}  % <== for grading, replace with your team name, e.g. q1-team-420
\date{Habib University | Spring 2023}

\qformat{{\large\bf \thequestion. \thequestiontitle}\hfill}
\boxedpoints

\begin{document}
\maketitle

\begin{questions}

  \titledquestion{Sacred Secrets}[5] One of your helpful TAs has prepared the manual, ``Sacred Secrets: How to Stay Sane and Earn an A+''. However their own sanity is sadly no longer intact. The \LaTeX\ source and the repository got deleted and all that exists about the location of the only printed copy are the following instructions.
  \begin{enumerate}
  \item There is a hint at Learn Courtyard or at the Gym.
  \item If your TA is sitting in Ehsas or they are absent, then there is a hint at Learn Courtyard.
  \item If your TA is not sitting in Ehsaas, then there is a hint at the Gym.
  \item If there are people in Learn Courtyard, then there is no hint at Learn Courtyard.
  \item If there is a hint at Learn Courtyard, then the manual is at Zen Garden.
  \item If there is hint at the Gym, then the manual is at Earth Courtyard.
  \item If your TA is absent, then the manual is at Fire Courtyard.
  \end{enumerate}
  You notice that there are people in Learn Courtyard. Show how you can infer the location of the manual.

  \begin{solution}\\
    % Enter your solution here.
    \textit{a = There is a hint at Learn Courtyard \\
    b = There is a hint at the Gym\\
    c = TA is sitting in Ehsas \\ d = TA is absent \\ e = There are people in Learn Courtyard \\ f = manual is at Zen Garden \\ g = manual is at Earth Courtyard \\ h = The manual is at Fire Courtyard} \\ 
    \\
    a \lor \ b \\ (c \lor \ d) \implies \ a 
    \\ \neg c \implies \ b \\ e \implies \neg a \\ a \implies \ f \\ b \implies \ g \\ d \implies \ h \\ 
    $Since, we know that there are people in Learn Courtyard so,
    e = True 
    then$ \neg a  \ $is also true according to,$
    \\ $e \implies \neg a$ \\
    this says that a is false, which lets b to be true according to, $a \lor \ b$ \\
    So, according to, \\ $b \implies \ g$ \\ the manual is at Earth Courtyard.

  \end{solution}

  \titledquestion{Logic Sauce}[5] In order to celebrate their performance in the recent quiz, a group of $n$ Discrete Mathematics students are hanging out at the Dhaba area. Rahim bhai asks them, ``Does everyone want fries?'' The first student replies, ``I do not know'', as does the second. In fact, the first $n-1$ students respond the same. The last student replies, ``No''. Rahim bhai then proceeds to prepare the order. What is the order and how did Rahim bhai deduce it?

  \begin{solution}
    % Enter your solution here.
    the first \textit{n-1} students will get the fries. As they answered to the question that asked if everyone wants fries, and they answered, 'I don't know' which means that they want fries, and they don't know about the remaining people. Whereas, the last person said No, which means that he does not want the fries.
  \end{solution}

  
\end{questions}

\end{document}


\documentclass[addpoints]{exam}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{geometry}
\usepackage{tabularx}
\usepackage{titling}

% Header and footer.
\pagestyle{headandfoot}
\runningheadrule
\runningfootrule
\runningheader{CS/MATH 113}{HW 2: Logic}{Solved by \theauthor}
\runningfooter{}{Page \thepage\ of \numpages}{}
\firstpageheader{}{}{}

\boxedpoints
\noprintanswers

\newcommand\ol\overline


\title{Homework 2: Logic}
\author{upper-bound}  % replace with your team name
\date{CS/MATH 113 Discrete Mathematics\\Habib University, Spring 2022}

\begin{document}
\maketitle

\begin{questions}

\section*{Propositional Logic}
  
\question Prove or disprove the following claims using truth tables. In each case, explicitly state your conclusion and how it is supported by the truth table.
  \begin{parts}
  \part[5] $\neg(p \lor q) \equiv \neg p \land \neg q $.
    \begin{solution}
      % Part of the table is given here for your convenience.
      \[
        \begin{array}{c|c|*{6}{|c}}
          p & q & p \lor q & \neg(p \lor q) & \neg p & \neg q & \neg p \land \neg q & (\neg(p \lor q)) \iff (\neg p \land \neg q)\\
          \hline
          F & F &  &  &  &  &  &  \\
          F & T &  &  &  &  &  &  \\
          T & F &  &  &  &  &  &  \\
          T & T &  &  &  &  &  &  
        \end{array}
      \]
      
    \end{solution}

  \part[5] $ (p \lor q) \implies \neg r \equiv (\neg p \land \neg q) \land \neg r$.
    \begin{solution}
      % Part of the table is given here for your convenience.
      For ease of notation, let
      \[
        \begin{array}{l@{\text{ : }} l}
          A & (p\lor q)\implies \lnot r \\
          B & (\lnot p \land \lnot q) \land \lnot r
        \end{array}
      \]
      So we have to prove that $A \equiv B$.
      \[
        \begin{array}{*{3}{c|}*{8}{|c}}
          p & q & r & \lnot p & \lnot q & \lnot r & p \lor q & \lnot p \land \lnot q & A & B & A \iff B\\
          \hline
          F & F &  &  &  &  &  &  &  &  &  \\
          F & F &  &  &  &  &  &  &  &  &  \\
          F & T &  &  &  &  &  &  &  &  &  \\
          F & T &  &  &  &  &  &  &  &  &  \\
          T & F &  &  &  &  &  &  &  &  &  \\
          T & F &  &  &  &  &  &  &  &  &  \\
          T & T &  &  &  &  &  &  &  &  &  \\
          T & T &  &  &  &  &  &  &  &  &  
        \end{array}
      \]
    \end{solution}
    
  \end{parts}

\question We want to write the statement, ``A person is popular only if they are cool or funny'', in propositional logic.
  \begin{parts}
  \part[5] Identify three simple propositions, $p, q, \text{ and } r$, needed for the representation and write out the corresponding expression that uses them to represent the given sentence.
    \begin{solution}
      % Enter your solution here.
    \end{solution}
  \part[5] For your expression identified above, write the converse, contrapositive, and inverse in propositional logic as well as complete English sentences.
    \begin{solution}
      
      % Part of the table is given here for your convenience.
      \begin{tabularx}{\textwidth}{l|l|X}
        & Logical Notation & English sentence \\\hline\hline
        Converse &   & \\\hline
        Contrapositive &  & \\\hline
        Inverse &  &  
      \end{tabularx}

    \end{solution}
  \end{parts}

\question[5] A small company makes widgets in a variety of constituent materials (aluminum, copper, iron), colors (red, green, blue, grey), and finishes (matte, textured, coated). Although there are many combinations of widget features, the company markets only a subset of the possible combinations. The following sentences are constraints that characterize the possibilities. 
  \begin{enumerate}
  \item aluminum $\lor$ copper $\lor$ iron
  \item aluminum $\implies$ grey
  \item copper $\land$ $\neg$ coated $\implies$ red
  \item coated $\land$ $\neg$ copper $\implies$ green
  \item green $\lor$ blue $\iff \neg$ textured $\land$ $\neg$ iron
  \end{enumerate}
  Suppose that a customer places an order for a copper widget that is both green and blue with a matte finish.
  \begin{parts}
  \part[5] Using the propositions above, express the order as a compound proposition in logical notation.
    \begin{solution}
      % Enter your solution here.
    \end{solution}
  \part[5] Determine which constraints are satisfied and which are violated for the order, and provide an explanation.
  \begin{solution}

    % Part of the table is given here for your convenience.
    \begin{tabularx}{\textwidth}{l|l|X}
      Constraint & Satisfied & Explanation \\\hline\hline
      aluminum $\lor$ copper $\lor$ iron & & \\
      aluminum $\implies$ grey &  &  \\\hline
      copper $\land$ $\neg$ coated $\implies$ red &  &  \\\hline
      coated $\land$ $\neg$ copper $\implies$ green &  &  \\\hline
      green $\lor$ blue $\iff \neg$ textured $\land$ $\neg$ iron &  & 
    \end{tabularx}
  \end{solution}

\end{parts}


\question[5] You are given four cards each of which has a number on one side and a letter on another. You place them on a table in front of you and the four cards read: $A\ 5\ 2\ J$. Which cards would you turn over in order to test the following rule? 
  \begin{center}
    Cards with $5$ on one side have $J$ on the other side.
  \end{center}
  Explain your choice.
  \begin{solution}

      % Part of the table is given here for your convenience.
    \begin{tabularx}{\textwidth}{c|c|X}
      Card & Turned & Explanation \\\hline\hline
      $A$ &  & \\\hline
      $5$ &  & \\\hline
      $2$ &  & \\\hline
      $J$ &  & 
    \end{tabularx}
    
  \end{solution}
  
\question An argument is said to be \textit{valid} if its \textit{conclusion} can be inferred from its \textit{premises}. An argument that is not valid is called an \textit{invalid} argument, or a \textit{fallacy}. For each of the arguments below, identify the simple propositions involved, write the premises and conclusion(s) in logical notation using the identified simple propositions, and decide whether it is valid. Justify your decision.

  \begin{parts}
  \part[5] If I am wealthy, then I am happy. I am happy, therefore, I am wealthy.
    \begin{solution}
      % Part of the structure is given here for your convenience.
      The simple propositions are as follows.\\
      \begin{tabularx}{\textwidth}{l@{ : }X}
        $p$ & \\ % state the atomic proposition
        $q$ & % state the atomic proposition
      \end{tabularx}

      The argument is
      \[
        \begin{array}{l}
          \text{enter the premise in logical notation}\\
          \text{enter the premise in logical notation}\\\hline
          \text{enter the conclusion in logical notation}\\
        \end{array}
      \]
    \end{solution}
  \part[5]
    If Ahmed drives his car, he is at least 18 years old. Ahmed does not drive a car. Therefore, Ahmed is not yet 18 years old. 
    \begin{solution}
      % Enter your solution here.
    \end{solution}
  \part[5] If I study, then I will not fail CS 113. If I do not play cards too often, then I will study. I failed CS 113. Therefore, I played cards too often.
    \begin{solution}
      % Enter your solution here.
    \end{solution}
  \end{parts}

  
\section*{Predicate Logic}
  
\question
  \begin{parts}
  \part[5] There is a third quantifier often used in predicate logic called the \textit{Uniqueness Quantifier}, $\exists!x\; P(x)$ which is read as, ``$P(x)$ is true for one and only one $x$ in the domain'', or ``there is a \textit{unique} $x$ such that $P(x)$''. Give an example of a propositional function $P(x)$ and a corresponding domain, such that $\exists!x\; P(x)$ is a true proposition.
    \begin{solution}
      % Enter your solution here.
    \end{solution}
    
  \part[5] The uniqueness quantifier can be expressed using the other two quantifiers but is still used on its own as it shortens the logical terms. In particular,
    \begin{align}
      \exists!x\;  P(x) \equiv \exists x\; (P(x) \land \forall y\; (P(y) \rightarrow y = x)) \label{eq:uniq}
    \end{align}
    Express the proposition on the right above in English and explain why it is equivalent to the left hand side, i.e. to the uniquely quantified propositional function. You may explain in words; a formal proof is not yet required.
    \begin{solution}
      % Enter your solution here.
    \end{solution}
    
  \part[5] Express $\neg \exists!xP\; (x)$ in a similar way as (\ref{eq:uniq}). Provide an expression in formal notation as well as in English. Also, give an example of a true proposition $\neg\exists!x\; P(x)$ by slightly changing the one you gave in part (a).
    \begin{solution}
      % Enter your solution here.
    \end{solution}
  \end{parts}

  
\question
  For each of the statements given below, perform the following.
  \begin{enumerate}
  \item Express the statement in formal notation using quantifiers.
  \item Express the negation of the statement in formal notation such that no negation is left to the quantifier.
  \item Express the negated statement above as a statement in English.
  \end{enumerate}

  \begin{parts}
  \part[5] No one can have Pakistani and Indian citizenship.
    \begin{solution}
      % Enter your solution here.
    \end{solution}

  \part[5] If everyone does their homework and goes to the recitations, no one will be badly prepared for the exams.
    \begin{solution}
      % Enter your solution here.
    \end{solution}


  \part[5] No student has solved at least one exercise in every section of the book.
    \begin{solution}
      % Enter your solution here.
    \end{solution}

    
  \part[5] No one has climbed every mountain in Pakistan.
    \begin{solution}
      % Enter your solution here.
    \end{solution}
  \end{parts}

\question
  Translate the specifications below into English using the given propositional functions.\\
  \begin{tabular}{l@{ : }l}
    $F(p)$ & The printer $p$ is out of service\\
    $B(p)$ & Printer $p$ is busy\\
    $L(j)$ & Print job $j$ is lost\\
    $Q(j)$ & Print job $j$ is queued
  \end{tabular}
  \begin{parts}
  \part[5] $\exists p\; (F(p) \land B(p)) \rightarrow \exists j\; L(j)$
    \begin{solution}
      % Enter your solution here.
    \end{solution}
    
  \part[5] $(\forall p\; B(p) \land \forall j\; Q(j)) \rightarrow \exists j\; L(j)$
    \begin{solution}
      % Enter your solution here.
    \end{solution}
  \end{parts}

\question Express each of the system specifications below using suitable predicates, quantifiers, and logical connectives.
  \begin{parts}
  \part[5] At least one mail message can be saved if there is a disk with more than 10KB of free space.
    \begin{solution}
      % Enter your solution here.
    \end{solution}

  \part[5] The system mailbox can be accessed by everyone in the group if the file system is locked.
    \begin{solution}
      % Enter your solution here.
    \end{solution}
  \end{parts}

\question
  Consider the propositions below for which the domain of all variables is $\mathbb{Z}$. For each proposition,
  \begin{enumerate}
  \item Express the proposition in English,
  \item State its truth value and provide an explanation if it is true or a counterexample if it is false, and
  \item Specify a domain for which the proposition has the other truth value.
  \end{enumerate}

  \begin{parts}
  \part[5] $\forall x \forall y\; (x^2= y^2 \rightarrow x=y)$
    \begin{solution}
      % Enter your solution here.
    \end{solution}

  \part[5] $\forall x \exists y\; (y^2=x)$
    \begin{solution}
      % Enter your solution here.
    \end{solution}

  \part[5] $\exists x \forall y\; (x \leq y^2)$
    \begin{solution}
      % Enter your solution here.
    \end{solution}

  \part[5] $\forall x \forall y\ \exists z\; (x-z=y)$
    \begin{solution}
      % Enter your solution here.
    \end{solution}
  \end{parts}
  
\end{questions}

\end{document}


%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
